---
title: "Exercise Sheet 3"
author: "Akeo Niedenzu (10537049)"
subtitle: "Fortgeschrittene Statistische Software für NF"
date: "`r Sys.Date()`"
output: distill::distill_article
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


## Exercise 1: Initializing git (4 Points)
For this whole exercise sheet we will be tracking all our changes to it in git.

### a)
Start by initializing a new R project with git support, called 2025-exeRcise-sheet-3. If you forgot how to do this, you can follow this guide.

- done


### b)
Commit the files generated by Rstudio.
 
- done


### c)
For all of the following tasks in this exercise sheet we ask you to always commit your changes after finishing each subtask e.g. create a commit after task 1d, 1e etc.
Note: This applies only to answers that have text or code as their answer. If you complete tasks in a different order or forget to commit one, this is no problem. If you change your answers you can just create multiple commits to track the changes

- done


### d)
Name 2 strengths and 2 weaknesses of git.

Strengths:
- good distributed version control - multiple developers can work on the same project with a full copy of the repository, including its history
- enables offline work and has good backup capabilities

Weaknesses:
- annoying setup / steep learning curve
- poor handling of large binary files (videos, datasets, etc.) can lead to bloat and slow down a repository


### e)
Knit this exercise sheet. Some new files will automatically be generated when knitting the sheet e.g. the HTML page. Ignore these files, as we only want to track the source files themselves. You can, but don't need to create a `.gitignore` file. Just do not commit these files manually.



## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

### a)
Create a new repository on GitHub in your account named `exeRcise-sheet-3`. Make sure you create a **public repository** so we are able to see it for grading. Add the link to the repository below:
https://github.com/ANiedenzu/exeRcise-sheet-3.git


### b)
Push your code to this new repository by copying and executing the snippet on github listed under `…or push an existing repository from the command line`.


### c)
Regularly push your latest changes to GitHub again and especially do so when you are finished with this sheet.



## Exercise 3: Pixar Films (4 Points)
Download the `pixar_films` and `public_response` datasets from the
GitHub repository and track them in git.
Link:
<https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-03-11>
For small datasets like these adding them to git is not a problem.

### a)
Load the `pixar_films` dataset into R. Clean the dataset by removing films without a title. Inspect the variable `film_rating`. What are the possible values and what do they mean? Create a factor variable for the film rating. Why is this appropriate?

```{r}
#load packages
library(tidyverse)

#load dataset
pixar_films = read_csv("data/pixar_films.csv")
#View(pixar_films)
```

Values for the variable "film_rating" are "G", "PG", "N/A" and "Not Rated". "N/A" and "Not Rated" referring to not (yet) existing ratings. "G" is short for "General Audiences" meaning the movie is suitable for all ages. "PG" is short for "Parental Guidance suggested" meaning the movie contains content that might not be suitable for young children. For example a PG-13 rated movie contains content inappropiate for pre teenage children. Parents are advised to accompany or guide children throughout the movie.

```{r}
#remove all movies with no titles, change film_rating to be a factor variable instead of a character variable, remove useless number column
pixar_films_ratingfactor = pixar_films %>%
  filter(!(is.na(film))) %>% 
  mutate(film_rating = factor(
    film_rating,
    levels = c("G", "PG", "N/A"),
    ordered = TRUE)
    ) %>%
  select(-number)


#View(pixar_films_ratingfactor)
#typeof(pixar_films_ratingfactor$film_rating)
#typeof(pixar_films$film_rating)
```


### b)
Inspect the film titles manually. Which films form a film series? A film series can be identified by a common word in the titles of the films, often in conjunction with a number in the title, e.g. "Despicable Me" and "Despicable Me 2". Create a dataframe which displays a list of the different series with the titles of the films and how many films belong to the series. Output the dataframe.

```{r}
#create "series" column giving all movies a series variable depending on specific strings in their title - "Standalone" for no match
pixar_films_series = pixar_films_ratingfactor %>%
  mutate(series = case_when(
    str_detect(film, regex("Toy Story", ignore_case = TRUE)) ~ "Toy Story",
    str_detect(film, regex("Cars", ignore_case = TRUE)) ~ "Cars",
    str_detect(film, regex("Finding", ignore_case = TRUE)) ~ "Finding Fish",
    str_detect(film, regex("Monsters", ignore_case = TRUE)) ~ "Monsters",
    str_detect(film, regex("Incredibles", ignore_case = TRUE)) ~ "Incredibles",
    TRUE ~ "Standalone")
    )

#create a new dataframe consisting of the series, their movies and the amount of movies within each series
#sort by amount of movies per series
#remove all standalone movies
series_summary = pixar_films_series %>%
  group_by(series) %>%
  summarise(
    films = paste(film, collapse = ", "),
    count = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(count)) %>% 
  filter(!(series=="Standalone"))

#show new dataframe
series_summary

```


### c)
Load the `public_response` dataframe into R. Convert the `cinema_score` variable into a factor while ensuring the factor levels are defined in ascending order, from the lowest to the highest score. Combine `public_response` with the `pixar_films` dataset using an appropriate merge variable.

```{r}
#load dataframe
public_response = read_csv("data/public_response.csv")
#View(public_response)
public_response_asfactor = public_response %>% 
  mutate(cinema_score = factor(
    cinema_score,
    levels = c("A-", "A", "A+", "NA"),
    ordered = TRUE
    )
  )

#View(public_response_asfactor)

pixar_films_public_response = left_join(x = pixar_films_ratingfactor, y = public_response_asfactor, by = "film")

#View(pixar_films_public_response)
```


### d)
Choose one of the variables representing the public response and create a bar plot for the films belonging to a series. Here are the details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.
    
    
```{r}
#group by series for combined dataframe
#remove standalone movies and order by release date within a series
pixar_combined_series = pixar_films_public_response %>%
  mutate(series = case_when(
    str_detect(film, regex("Toy Story", ignore_case = TRUE)) ~ "Toy Story",
    str_detect(film, regex("Cars", ignore_case = TRUE)) ~ "Cars",
    str_detect(film, regex("Finding", ignore_case = TRUE)) ~ "Finding Fish",
    str_detect(film, regex("Monsters", ignore_case = TRUE)) ~ "Monsters",
    str_detect(film, regex("Incredibles", ignore_case = TRUE)) ~ "Incredibles",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(series)) %>%
  arrange(series, release_date) %>%
  mutate(film = factor(film, levels = unique(film)))

View(pixar_combined_series)

#plot using audience_score
ggplot(pixar_combined_series, aes(x = series, y = metacritic, fill = film)) +
  geom_bar(
    stat = "identity",
    width = 0.8,
    position = position_dodge(preserve = "single", width = 1)
    ) +
  labs(
    title = "Metacritic Scores for Pixar Films",
    subtitle = "grouped by series",
    x = "Film Series",
    y = "Metacritic Score",
    fill = "Film"
    ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?

The first film always has the best audience score in each series with the second film always producing a noticeable drop. Toy Story seems to be generally the most likes series, which is why it's not surprising it has the most amount of films of all the series.



## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

### a)
Go to <https://github.com/owid/owid-datasets/tree/master/datasets> and choose a dataset that interests you. You can have a look at <https://ourworldindata.org/> to gather some inspiration.

- done


### b)
Download the dataset and track it in git.
```{r}
roaddeaths = read_csv("data/German road deaths and accidents - destatis.csv")
```


### c)
Put the name / title of the dataset and a link to it below.

-   Dataset Name: German road deaths and accidents - destatis
-   Link: <https://github.com/owid/owid-datasets/blob/master/datasets/German%20road%20deaths%20and%20accidents%20-%20destatis/German%20road%20deaths%20and%20accidents%20-%20destatis.csv>


### d)
Come up with a (research) question yu want to answer with the data and briefly explain why you believe this is an interesting question within one sentence. It should be a question that can be answered with the dataset and using R.


### e)
Use R to answer your chosen question. Please limit your analysis to the functions and techniques we have covered so far in the course. You are **not expected** to use advanced statistical models or external packages which haven't been introduced.


### f)
Create a meaningful plot / figure with the dataset. Make sure to provide a figure caption (via the chunk options / Rmarkdown) and correctly label the figure.



## Final Note
Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.